---
title: "Palmer Penguins"
output:
  html_document: 
    keep_md: yes
    theme: cerulean
  pdf_document: default
  word_document: default
 
---
<b>Source:</b> [Dr. Kristen Gorman, Dr. Allison Horst, and Dr. Alison Hill.](https://doi.org/10.1371/journal.pone.0090081) <br>
[Data](https://allisonhorst.github.io/palmerpenguins/)

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE ,warning = F,error = F,message = F,
                      fig.align = "center")
                      
```



### Import the necessary packages

```{r}
{ 
  library(tidyverse)
  library(caret)
  library(plotly)
  library(data.table)
  library(psych)
  library(hrbrthemes)
  library(corrr)
  library(rpart)
  library(rpart.plot)
  library(cvms)

}

```



### Import the data

```{r}
# tuesdata <- tidytuesdayR::tt_load(2020, week = 31)
# penguins <- tuesdata$penguins
# penguins_raw<- tuesdata$penguins_raw
penguins<-read.csv("penguins.csv",header = T,na.strings = c(""," ","NA"))
penguins_raw<-read.csv("penguins_raw.csv",header = T,na.strings = c(""," ","NA"))

```

### Lets check the structure and summary of the data


```{r}
str(penguins)
summary(penguins)
```


### Lets check the various Species of Penguins and its counts

```{r }
penguins %>% group_by(species) %>% summarise(counts = n()) %>%
  plot_ly(
  labels = ~ species,
  values = ~ counts,
  type = 'pie',
  hole = 0.5,
  sort = F,
  textinfo = "label",
  direction = "clockwise",
  title = "<b>Species of Penguins</b>",
  showlegend = F
  )
```

### Lets visualize various Species of Penguins by Sex and Island

```{r,fig.width = 10,fig.height =8}
penguins  %>%  
  group_by(species,sex,island) %>% 
  summarise(counts=n()) %>% 
  ggplot(aes(x = "", y = counts, fill = species )) + 
  geom_bar(stat = "identity", position = position_fill()) +
  geom_text(aes(label = counts), size=5,
            fontface="bold",color="white",
            position = position_fill(vjust = 0.5)) +
  scale_fill_manual(values=c("darkorange", "orchid", "darkcyan"))+
  guides(fill=guide_legend(nrow=1, title.position = "top"))+
  coord_polar(theta = "y") +
  facet_grid(~sex~island)  +
  theme(axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        strip.background.x =   element_rect(fill = "gold2"),
        strip.background.y =   element_rect(fill = "thistle2"),
        strip.text.x = element_text(size = 10, face="bold"),
        strip.text.y = element_text(size = 10, face="bold"),
        legend.position="bottom",
        legend.title = element_text(size = 10, face="bold",hjust = 0.5),
        legend.text = element_text(size = 9, face="bold"),
        axis.text = element_blank(),
        axis.ticks = element_blank(),
        panel.grid  = element_blank())
```

#### The above pie chart explains that in the Torgersen island only adelie penguins were identified, in that 5 species sex were not identified.In Dream island Adelie and Chinstrap penguins were identified.Gentoo penguins were only identified in Biscoe island.

### Lets perform the Exploratory Data Analysis


### Compare the bill_length_mm distribution with species and compare them with Sex and Island

#### Create a function to get the median value as dataframe

```{r}
fun_median <- function(x){
  return(data.frame(y=median(x),label=round(mean(x,na.rm=T) ,0) ))}
```
#### Import the flat violin funtion from github url

```{r}
source(
  "https://gist.githubusercontent.com/benmarwick/2a1bb0133ff568cbe28d/raw/fb53bd97121f7f9ce947837ef1a4c65a73bffb3f/geom_flat_violin.R"
)
```


```{r,fig.width = 10,fig.height =8}

penguins %>% drop_na()    %>% 
  ggplot(aes(y=bill_length_mm,x=species,fill= species)) +
  geom_flat_violin(position = position_nudge(x = .2, y = 0), alpha = .8)+
  geom_point(aes(y = bill_length_mm, color = species), 
             position = position_jitter(width = .15), size = .5, 
             alpha = 5)+
  geom_boxplot(width=.1,
               outlier.fill  ="red", outlier.shape=23,
               outlier.size=2,inherit.aes = T
  ) +scale_fill_manual(values=c("darkorange", "orchid", "darkcyan"))+
  scale_color_manual(values=c("darkorange", "orchid", "darkcyan"))+
  stat_summary(fun = median, geom="point",colour="darkred", size=3,
               show.legend = FALSE) +coord_flip()+
  stat_summary(fun.data = fun_median,color="black", geom="text", vjust=-0.7)+
  facet_grid(~sex~island)+
  theme(  
    legend.position = "none",
    strip.background.x =   element_rect(fill = "gold2"),
    strip.background.y =   element_rect(fill = "thistle2"),
    strip.text.x = element_text(size = 10, face="bold"),
    strip.text.y = element_text(size = 10, face="bold"),
    axis.title.x = element_text(face="bold",hjust = 0.5),
    axis.title.y = element_text(face="bold",vjust = 0.5),
    axis.text.y = element_text(face="bold"),
    axis.text.x = element_text(face="bold"),
    axis.line.x = element_blank(),
    axis.line.y = element_blank(),
    axis.ticks = element_blank())  
```


#### Outliers("Red Diamond")


### Compare the bill_depth_mm distribution with species and compare them with Sex and Island


```{r,,fig.width = 10,fig.height =8}

penguins %>% drop_na()    %>% 
  ggplot(aes(y=bill_depth_mm,x=species,fill= species)) +
  geom_flat_violin(position = position_nudge(x = .2, y = 0), alpha = .8)+
  geom_point(aes(y = bill_depth_mm, color = species), 
             position = position_jitter(width = .15), size = .5, 
             alpha = 5)+
  geom_boxplot(width=.1,
               outlier.fill  ="red", outlier.shape=23,
               outlier.size=2,inherit.aes = T
  ) +scale_fill_manual(values=c("darkorange", "orchid", "darkcyan"))+
  scale_color_manual(values=c("darkorange", "orchid", "darkcyan"))+
  stat_summary(fun = median, geom="point",colour="darkred", size=3,
               show.legend = FALSE) +coord_flip()+
  stat_summary(fun.data = fun_median,color="black", geom="text", vjust=-0.7)+
  facet_grid(~sex~island)+
  theme(  
    legend.position = "none",
    strip.background.x =   element_rect(fill = "gold2"),
    strip.background.y =   element_rect(fill = "thistle2"),
    strip.text.x = element_text(size = 10, face="bold"),
    strip.text.y = element_text(size = 10, face="bold"),
    axis.title.x = element_text(face="bold",hjust = 0.5),
    axis.title.y = element_text(face="bold",vjust = 0.5),
    axis.text.y = element_text(face="bold"),
    axis.text.x = element_text(face="bold"),
    axis.line.x = element_blank(),
    axis.line.y = element_blank(),
    axis.ticks = element_blank())  
```



#### Outliers("Red Diamond").


### Lets analysis the relation between bill_length_mm and bill_depth_mm and compare them with species of penguins.

```{r,fig.width = 10,fig.height = 8}
penguins %>%  ggplot(aes(x = bill_length_mm, y = bill_depth_mm, col = species)) +
        geom_smooth(method = "lm", se = T) +
        scale_color_manual(values=c("darkorange", "orchid", "darkcyan"))+
        geom_jitter() +
        coord_equal()+
   guides(fill=guide_legend(nrow=1, title.position = "top"))+

  theme(
        panel.background = element_blank(), 
        legend.position = "bottom",
        legend.title = element_text(size = 10, face="bold",hjust = 0.5),
        legend.text = element_text(size = 9, face="bold"),
        strip.background.x =   element_rect(fill = "gold2"),
        strip.text.x = element_text(size = 10, face="bold"),
        axis.title.x = element_text(face = 'bold', size = 12),
        axis.title.y = element_text(face = 'bold', size = 12),
        axis.text.x = element_text(face = 'bold', size = 11),
        axis.text.y = element_text(face = 'bold', size = 11)
       )
  
```

```{r}
data.table(penguins %>% drop_na() %>%  group_by(species)  %>% 
dplyr::summarize(cor(bill_length_mm,bill_depth_mm)))
```


#### The above analysis explains Chinstrap and Gentoo species having high correlation on bill_length_mm and bill_depth_mm when compare to Adelie Species.


### Compare the flipper_length_mm distribution with species and compare them with Sex and Island.

```{r,fig.width = 10,fig.height = 8}
penguins %>% drop_na()    %>% 
  ggplot(aes(y=flipper_length_mm,x=species,fill= species)) +
  geom_flat_violin(position = position_nudge(x = .2, y = 0), alpha = .8)+
  geom_point(aes(y = flipper_length_mm, color = species), 
             position = position_jitter(width = .15), size = .5, 
             alpha = 5)+
  geom_boxplot(width=.1,
               outlier.fill  ="red", outlier.shape=23,
               outlier.size=2,inherit.aes = T
  ) +scale_fill_manual(values=c("darkorange", "orchid", "darkcyan"))+
  scale_color_manual(values=c("darkorange", "orchid", "darkcyan"))+
  stat_summary(fun = median, geom="point",colour="darkred", size=3,
               show.legend = FALSE) +coord_flip()+
  stat_summary(fun.data = fun_median,color="black", geom="text", vjust=-0.7)+
  facet_grid(~sex~island)+
  theme(  
    legend.position = "none",
    strip.background.x =   element_rect(fill = "gold2"),
    strip.background.y =   element_rect(fill = "thistle2"),
    strip.text.x = element_text(size = 10, face="bold"),
    strip.text.y = element_text(size = 10, face="bold"),
    axis.title.x = element_text(face="bold",hjust = 0.5),
    axis.title.y = element_text(face="bold",vjust = 0.5),
    axis.text.y = element_text(face="bold"),
    axis.text.x = element_text(face="bold"),
    axis.line.x = element_blank(),
    axis.line.y = element_blank(),
    axis.ticks = element_blank())  
  
```




#### Outliers("Red Diamond")  



### Compare the body_mass_g distribution with species and compare them with Sex and Island.

```{r,fig.width = 10,fig.height = 8}
penguins %>% drop_na()    %>% 
  ggplot(aes(y=body_mass_g,x=species,fill= species)) +
  geom_flat_violin(position = position_nudge(x = .2, y = 0), alpha = .8)+
  geom_point(aes(y = body_mass_g, color = species), 
             position = position_jitter(width = .15), size = .5, 
             alpha = 5)+
  geom_boxplot(width=.1,
               outlier.fill  ="red", outlier.shape=23,
               outlier.size=2,inherit.aes = T
  ) +scale_fill_manual(values=c("darkorange", "orchid", "darkcyan"))+
  scale_color_manual(values=c("darkorange", "orchid", "darkcyan"))+
  stat_summary(fun = median, geom="point",colour="darkred", size=3,
               show.legend = FALSE) +coord_flip()+
  stat_summary(fun.data = fun_median,color="black", geom="text", vjust=-0.7)+
  facet_grid(~sex~island)+
  theme(  
    legend.position = "none",
    strip.background.x =   element_rect(fill = "gold2"),
    strip.background.y =   element_rect(fill = "thistle2"),
    strip.text.x = element_text(size = 10, face="bold"),
    strip.text.y = element_text(size = 10, face="bold"),
    axis.title.x = element_text(face="bold",hjust = 0.5),
    axis.title.y = element_text(face="bold",vjust = 0.5),
    axis.text.y = element_text(face="bold"),
    axis.text.x = element_text(face="bold"),
    axis.line.x = element_blank(),
    axis.line.y = element_blank(),
    axis.ticks = element_blank())     
  
```


#### Outliers("Red Diamond")  


### Lets find out the correlation between the continous variables and visualize it by using network plot.

```{r,fig.width = 10,fig.height =10,units="cm",dpi=300 }

res.cor <- correlate(penguins %>%select(-year) %>% 
                       drop_na() %>% select_if(is.numeric),
                     method="spearman",diagonal=0) 
 
cornetplot <- network_plot(
  res.cor ,
  colors = c("indianred2", "white", "dodgerblue"),
  min_cor = .5,
  repel = T,
  curved = T
  )
  
print(
  cornetplot + ggplot2::labs(
  title = "Correlations between Bill Length, Bill Depth, Flipper Length, Body Mass",
  subtitle = "Palmer Penguins",
  caption = "Source: Dr. Kristen Gorman, Dr. Allison Horst, and Dr. Alison Hill"
  )
  )
  
  
  
     
```

####  The above network plot shows that there is a high positive correlation between body_mass_g and flipper_length_mm. flipper_length_mm and bill_depth are negatively correlated.



### Lets split the data and perform tree model and evaluate the model.


```{r}
## convert character variables into factor

penguins<-penguins %>% mutate_if(is.character,as.factor)

penguins<-penguins %>% na.omit()

## split the train and validation data
set.seed(487)
index<-createDataPartition(y=penguins$species, p=0.75,list = F)
traind<-penguins[index,]
vald<-penguins[-index,]


## Build tree model 
treemodel<-rpart(species~. ,data = traind[,c(1,3:6)])




rpart.plot(treemodel,snip = T)


modpre<-predict(treemodel,vald,type = "class")

## Lets create the confusion matrix
con_mat<- confusion_matrix (modpre, vald$species)
con_mat

```

### Confusion Matrixplot

```{r}
plot_confusion_matrix(con_mat$`Confusion Matrix`[[1]])

```


